Group Ten
Abby Vece, Nina Braddock, Evan Satterfield

Graph ideas
- race of officer vs race of complainer
- number of complaints per rank
- number of complaints per month
- confidence interval for how long the case took vs board disposition 
- male vs female officer proportions of board disposition

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center")
pacman::p_load(tidyverse)

complaints <- read.csv("C_officercomplaints allegations_202007271729(2).csv")
```
intro
  This data is a list of complaints against the New York City police department. it was created in July 2020. The source is ProPublica. This data has reports on officers that are still in the force as of June 2020, so there are more complaints of officers who aren't on the force that aren't included in this data set, so if these officers were fired because of their actions, they aren't included, which could entirely change the data. All the reports were investigated by the CCRB, who were not able to come to any conclusions in many of the cases because the officers weren't very forthcoming with their evidence (such as body cam footage), so this data is probably not even close to what the actual data should be, but there is still a great amount of information to analyize. This data is important because it came in the wake of the George Floyd case, so people were very critical of police officers and their actions, so having access to information like this can really give good insight into what police officers are doing, especially in a city as big as New York City. Any data cleaning for this set what done for each question individually. 



```{r case vs board}
# confidence interval for how long the case took vs board disposition

complaints_intervals <-
  complaints |>
  filter(board_disposition != "Substantiated (MOS Unidentified)") |> # deleted cuz there was only one instance
  select(unique_mos_id, month_received, year_received, month_closed, year_closed, board_disposition) |> 
  mutate(months_takes = ((year_closed - year_received) * 12) + (month_closed-month_received)) |> 
  group_by(board_disposition) |> 
  summarize(num_complaints = n(),
            median = median(months_takes),
            average = mean(months_takes),
            month_sd = sd(months_takes)) |>
  mutate(t = qt(p = 0.975, df = n() -1),
         lower = average - t * (month_sd/sqrt(num_complaints)),
         upper = average + t * (month_sd/sqrt(num_complaints)),
         lower2 = case_when(lower < 0 ~ 0,
                            lower >= 0 ~ lower))

ggplot(data = complaints_intervals,
       mapping = aes(x = average,
                     y = fct_reorder(board_disposition, average)))+
  geom_segment(mapping = aes(x = lower2,
                             xend = upper,
                             y = board_disposition,
                             yend = board_disposition),
               color = "darkgrey")+
  geom_point() +
  theme_bw()+
  labs(x = "Average number of months for case to clear",
       y = "Board Disposition")+
  theme(axis.ticks = element_blank())+
  scale_x_continuous(minor_breaks = NULL)+
  expand_limits(x = 0) +
  geom_text(mapping = aes(label = paste0("n = ", num_complaints),
                          vjust = -.7))

```
  This graph shows the average number of months cases take to close. most of the cases take between 7 and 11 months to close, which is quite a long time. These cases were probably dragged on by the officers in hopes of the complainants wanting to drop the case. The quickest cases to close are the substantiated with no recommendations. This is probably because these are the cases that police officers want to close, and they took blame to get the case to blame quickly, knowing there were going to be no repercussions. The substantiated cases that do have repercussions also take a longer time, most likely because the police officer did something worse and were actively fighting it. One of the board dispositions was removed because there was only one case of it happening, so a confidence interval couldn't be created.  


```{r complaint and gender}
#box and whisker of age of complainant, gender, and officer gender
complaints_gend <-
  complaints |> 
  filter(complainant_age_incident > 0) |> 
  select(unique_mos_id, complainant_age_incident,complainant_gender, mos_gender) |> 
  mutate(complainant_gender_clean = case_when(complainant_gender == "Female" ~ "Female",
                                              complainant_gender == "Male" ~ "Male",
                                              TRUE ~ "Other")) |> 
  mutate(mos_gender = case_when(mos_gender == "F" ~"Female",
                                mos_gender== "M" ~ "Male"))

ggplot(data = complaints_gend,
       mapping = aes(x = complainant_gender_clean,
                     y = complainant_age_incident))+
  facet_wrap(~mos_gender)+
  geom_boxplot()+
  theme_bw() +
  labs(y = "Complainant age at Incident",
       x = "Complainant Gender")+
  scale_y_continuous(expand = c(0,0,0.5,0))

```
  This graph shows the age and gender of the complainant on the X and Y axis, in regard to the gender of the officer complained about. My original thought when making this graph would be that male officers would have more complaints from younger women in terms of sexual harassment, but that is not the case. Male officers tend to have more complaints from younger men. Female officers tend to get complaints evenly from everyone. The other category in this table includes transgender people, nonbinary people, and people who did not include their gender in their complaint, so this could change the data people had included their genders. 


```{r complaints per month}
complaints_month <-
  complaints |> 
  select(month_received) |> 
  count(month_received, name = "count")

ggplot(data = complaints_month, mapping = aes(x = month_received, y = count)) +
  theme_classic() +
  scale_y_continuous(breaks=seq(0, 4000, by = 1000)) +
  scale_x_continuous(breaks=seq(0, 12, by = 1)) + #expand = c(0,0,0.05,0)
  geom_col(fill = "steelblue", color = "gray20") +
  ggtitle("Complaints Received by Month") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Month") + 
  ylab("Number of Complaints Received")
```
  This graph shows the number of complaints received per month. The x axis is the month the complaint was received and the y axis is how many were received in that month. This graph ignores the year that the complaint was received. The goal of this graph was to see if the time of year affects the amount of complaints received. It seems like complaints are slightly more common during the spring and fall and noticeably lower during December, possibly due to the holidays.
  
```{r complaints per rank}
complaints_rank <-
  complaints |> 
  select(rank_incident) |> 
  count(rank_incident, name = "count")

complaints_rank$rank_incident[2] <- "Chief"
complaints_rank$rank_incident <- factor(x = complaints_rank$rank_incident,
                                        levels = c("Police Officer", "Detective", "Deputy Inspector", 
                                                   "Inspector", "Sergeant", "Lieutenant", "Captain", "Chief"))
ggplot(data = complaints_rank, 
       mapping = aes(x = rank_incident, y = count)) +
  theme_classic() +
  geom_col(fill = "steelblue", color = "gray20") +
  ggtitle("Complaints Received by Rank") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Rank") + 
  ylab("Number of Complaints Received") 
  # scale_x_discrete(expand = c(0,0,0.05,0))
```
  This graph shows the number of complaints received per rank of the officer. The x axis is the rank of the officer and the y axis is the number of complaints. This graph was made to see if certain ranks disproportionately receive complaints, and as expected the ranks that have to interact with people receive significantly more complaints than the ones that don't have to as much. 
  
  
```{r, race of officer vs race of complainer}
complaints |> 
  filter(allegation %in% c("Vehicle", "Physical force", "Gun Pointed", "Property damaged", "Chokehold")) ->
  five

ggplot(data = five) +
  geom_bar(mapping = aes(x = allegation,
                     fill = mos_ethnicity),
       position = "fill",
       color = "gray12",
       linewidth = 0.2) +
  theme_minimal()+
  scale_y_continuous(expand = c(0, 0, 0.05, 0),
 labels = scales::percent,
 breaks = (0:10)/10) +
  labs(x = "Allegations",
       y = "Proportion",
       fill = "Ethnicity") 
```
This graph illustrates five main allegations found in the data set as proportional bar graphs. Each bar
additionally looks at the complainant's ethnicity for that specific allegation. By setting position equal to fill, 
the individual allegations can be looked at on their own an compare each ethnicity to them respectfully or be used
to compare across allegations. For example, the proportion of allegations made against Hispanic individuals is greater
for vehicles than any of the other four selected allegations. In general, there are more allegations made towards 
white ethnicity because they represent a larger proportion of the sampled population. This also means that since
American Indian's make up a small portion of the general population that the sample size of this study has the possibility
of not including that group as, seen in this sample of allegations. While omitting this group from the data would not 
change the visual appearance of the graph it is still important to include American Indian's because their representation
in the data provides better context for where this study was done itself. 

```{r, complainant_age_incident vs years}
five |> 
  select(complainant_age_incident) |> 
  mutate(age_category = factor(case_when(
    complainant_age_incident >= 50 ~ "50+",
    complainant_age_incident >= 40 & complainant_age_incident < 50 ~ "40-49",
    complainant_age_incident >= 30 & complainant_age_incident < 40 ~ "30-39",
    complainant_age_incident >= 25 & complainant_age_incident < 30 ~ "25-29",
    complainant_age_incident >= 20 & complainant_age_incident < 25 ~ "20-24",
    complainant_age_incident >= 18 & complainant_age_incident < 20 ~ "18-19",
    complainant_age_incident < 18 ~ "under 18",
  ), levels = c("under 18", "18-19", "20-24", "25-29", "30-39", "40-49", "50+", "NA")))|> 
  ggplot(mapping = aes(x = age_category)) + 
  geom_bar(fill = "seagreen3",
           color = "darkgreen") +
  scale_y_continuous(breaks=seq(0, 2000, by = 200),
                     expand = c(0,0,0.05,0)) +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title = "Complaintant Age on the Total Number of Complaints Made",
         x = "Age",
         y = "Counts") 
```

This bar graph shows the relationship between different ages and the total number of complaints made by each of these groups.
As many would anticipate, the cluster of age groups makes a bell curve shaped graph. More specifically, the graph is right-skewed (positive skew)
which may be influenced by the younger age groups being more mobile and put into situations where they need to call the police. Personally,
I was expecting the data to be skewed slightly left because of the assumption that older people would understand the criminal justice system 
better and want to report more wrong doings. It could be the case that the younger age groups just have more interactions overall with the police. This would mean that younger people would have a larger representation in the sample size of this study which would cause the graph to 
be right-skewed. 

```{r}
#classification tree
#using officer rank and gender, complainant race, and fado_type to predict outcome

complaints_tree <-
  complaints |> 
  mutate(board_disposition_clean = case_when(board_disposition == "Exonerated" ~"Exonerates",
                                             board_disposition == "Unsubstantiated" ~ "Unsubstantiated",
                                             TRUE ~ "Substantiated"),
         months_takes = ((year_closed - year_received) * 12) + (month_closed-month_received),
         complainant_ethnicity_clean = case_when(complainant_ethnicity == "American Indian"~ "American Indian",
                                                 complainant_ethnicity == "Asian"~ "Asian",
                                                 complainant_ethnicity == "Black"~ "Black",
                                                 complainant_ethnicity == "Hispanic"~ "Hispanic",
                                                 complainant_ethnicity == "Other Race"~ "Other Race",
                                                 complainant_ethnicity == "Refused"~ "Refused",
                                                 complainant_ethnicity == "Unknown"~ "Unknown",
                                                 complainant_ethnicity == "White"~ "White",
                                                 TRUE ~ "Unknown"))

fado_tree_full <-
  rpart(formula = board_disposition_clean ~ fado_type  +rank_incident + mos_gender + complainant_ethnicity_clean,
        data = complaints_tree,
        method = "class",
        parms = list(split = "information"),
        minsplit = 2,
        minbucket = 1,
        cp = -1)

 rpart.plot(x = fado_tree_full,
             type = 5,
             extra = 101)

fado_tree_full$cptable |> 
  data.frame() |> 
  slice_min(xerror, n = 1) |> 
  mutate(xcutoff = xerror +xstd) |> 
  pull(xcutoff) ->
  xcutoff

fado_tree_full$cptable |> 
  data.frame() |> 
  filter(xerror < xcutoff) |> 
  slice(1) |> 
  pull(CP)->
  cp_prune

fado_prune<-
  prune(tree = fado_tree_full,
        cp = cp_prune)


rpart.plot(x = fado_prune,
           type = 5,
           extra = 101,
           box.palette = "pink")
```
For machine learning, we made a classification tree to try and predict the outcome of complaints, whether they were substantiated, unsubstantiated, or exonerated. We went through many different trials of adding and taking away different variables, until we could get the best tree possible. The variables we used were fado type, complainant ethnicity, the rank of the officer at the time of the incident, and the officer's gender. As shows previously, there were more than three outcomes, but for the sake of understanding, we combined the different substantiated groups into just one outcome. The classification tree only has one leaf for substantiated. The substantiated complaints are complaints that were looked into and the officer was found guilty of breaking NYPD rules. The exonerated complaints are complaints that were looked into and the officer was found guilty, but what they did broke no NYPD rules, and Unsubstantiated complaints are complaints where the officer was not found guilty. Only 3% of the data is sorted into the substantiated outcome in the classification tree, despite almost 25% of outcomes being substantiated, and the tree as a whole was only correct about 50% of the time, so the classification tree isn't a hugely accurate in predicting outcomes. We chose the classification tree because the variable we were trying to predict was categorical and we were using categorical explanatory variables. 
